Is malaria control profitable? Return on investment of privately-managed residential

### 2018-05-11 inst/rmd/01_introduction.Rmd
changes
removed text

Malaria accounts for a half million annual deaths worldwide [@White2014] [@World2016] [@Ashley2018]. Though rapid improvements in technology and funding have lead to important reductions in marlaria's global burden, the scale-up in activities required for eradication ("the worldwide interruption of transmission") [@TheLancet2011] will mean new partnerships and actors. One promising - albeit atypical - potential stakeholder in global malaria eradication is the private sector, given its omnipresence and potential to benefit directly from the elimination of malaria. But little evidence exists demonstrating how private sector entities can engage with malaria control and benefit at the firm level.

At the societal level, malaria has a large economic impact. By affecting saving, investment [@Shretta2016], risk perception, productivity, absenteeism [@Nonvignon_2016], human capital accumulation [@CastelBranco2014], mortality, and costs of care [@Sachs2002], malaria likely has a negative effect on GDP and growth [@McCarthy_2000] [@Orem_2012]. Because of the relative affordability of most intereventions and the enormous societal costs of malaria, most forms of malaria control are cost-effective when a public welfare perspective is assumed, such as when a government provides the financing [@White_2011] [@Purdy_2013] [@Howard_2017].

From the persepctive of the private sector, however, investing in malaria control is not so clear-cut. Public health interventions targetting malaria - and their corresponding cost-effectiveness evaluations - most often focus on impacts pertaining to public welfare, such as an increase in life years adjusted for disability or quality  [@Goodman1999] [@Shretta2016] [@Lee2017] [@Hanson2004]. Though population-level health is certainly of importance to businesses, and improvements in health incidentally improve the economy at all levels [@World1999] [@Bloom2008] [@Vecchi_2013], these improvements may be too disperse or long-term to incentivize private sector involvement in health campaigns.

100% of the Mozambican population are at risk of malaria, living in what the WHO classifies as a "high transmission" area [@Moonasar_2016]. Annually, Mozambique has more than 8 million clinical malaria cases (an annual incidence of approximately 300 per 1,000 residents), with an estimated 14,000 deaths. Malaria accounts for 29% of all deaths, and 42% of deaths among those under five years of age [@estatistica2009]. Since 2013, Mozambique has seen a gradual increase in the incidence of malaria [@Moonasar_2016]. 100% of the malaria in Mozambique is of the Plasmodium falciparum species, with Anopheles funestus, gambiae, and arabiensis as the primary mosquito vectors of the disease [@whoprof].

A significant sector of the economy in Mozambique is dominated by large-scale foreign direct investment projects [@Robbins2012], and the role of the private sector in health generally, and malaria specifically, is unequivocally important. Large agriculture and extractive industry firms take up wide swaths of land and employ hundreds of thousands  [@German2013]. The Mozambican state has encouraged large-scale entreprise with the aim of general economic development [@Buur2012]. And where large firms exist, they often take on social roles such as housing and health care [@Winkler]. At times, this role is necessary from a purely practical standpoint; in other cases, it is employed under the auspices of "corporate social responsibility" [@Azemar2009]. Regardless of the language used, it is clear that private industry plays an important role in public health in Mozambique [@Robbins2012] [@CastelBranco2014].

Several cases exist of foreign firms engaging in large-scale malaria control campaigns [@Mouzin2011] [@Han] [@Bennett_2017] [@Kaula_2017]. But these studies generally consider population health as the outcome measure of interest, rather than worker absenteeism or productivity. Similarly, they often neglect to differentiate between those clinical costs which are absorbed by the local health system versus those which are absorbed by the firm itself. In the literature, making the "investment case" for malaria control or elimination generally implies that the investor is the public sector, and takes into account those costs and benefits which are applicable from a public welfare point of view [@Shretta_2017]; though appropriate in most cases (the government or institutions interested in public welfare primarily being the primary malaria control agents in most locations), the findings of these studies are rarely applicable to the private sector, and even less so at granular levels (such as an individual firm). In the case of a private firm not interested in "corporate social responsibility", it is not clear whether investing in malaria control would be profitable or not. This lack of clarity not only discourages investment, but also makes it difficult for governments to pinpoint the correct of amount of subsidy (if applicable) to encourage private sector scale-up in malaria control.

To address the question of the profitability of malaria control activities from the standpoint of a private firm, we analyze data during a 4 year period from a private sugar facility in Southern Mozambique. We assess the effect of indoor residual spraying (IRS) on the absenteeism and health of workers, and demonstrate that the firm's engagement in malaria control not only improved worker health, but also generated a positive return on investment from a pure accounting perspective.

The structure of the paper is as follows. After a brief review of the existing literature on the subject of private sector investment in malaria control and its effects on worker productivity, we provide an overview of the sugar company under study, and the epidemiology of malaria in the nearby area, as well as in Mozambique as a whole. We then give an overview of the data collected, and outline the theoretical and methdological assumptions and tools that underly our analysis. In the results, we show the effectiveness of the company's malaria control program in reducing absences, and translate this reduction into cost savings. Our discussion covers potential implications from this study in terms of policy and investment, as well as the paper's limitations.

The literature on the effect of sugarcane cultivation on malaria risk is mixed. While some studies have found that the prevalence of malaria vectors in sugarcane areas to be similar to that of uncultivated areas (and less than in areas dedicated to other forms of more water-inensive agriculture, such as rice) [@Ijumba2002], other studies have found significant increases in factors associated with malaria transmission at large-scale sugarcane facilities relative to traditional, small-scale farming and non-irrigated farming [@Jaleta2013]. Regardless of the effect the presence of a sugarcane farm _per se_ on local malaria epidemiology, the time spent outdoors by sugarcane workers, the fact that many workers are migrants, and their sometimes precarious housing situations, suggest that sugarcane farmers are likely at increased risk of malaria infection [@OLaughlin2016]. This is important, given that even among occupations with far less inherent exposure to mosquitoes (such as health professionals), malaria is one of the primary causes of work absenteeism in malaria-endemic countries [@Burton1999]. There is also some concern regarding the effect of large-scale insecticide use - common at essentially all Sub-Saharan African sugarcane farms - on insecticide resistance among moquitoes in the area. A study in Belize found that mosquito populations on the edge of sugarcane fields had higher tolerance to insecticide than similar populations in the core of fields or outside of the periphery [@Dusfour2009]. Sugarcane areas may offer the standing water necessary for mosquito breeding, but also perhaps attract mosquitoes which would otherwise be elsewhere, due to compounds in sugarcane pollen [@Wondwosen2018].

Regardless of the academic literature on the risk of workers becoming infected with malaria, the companies running large agricultural operations in malaria-endemic regions generally consider the disease to be an important enough threat to workers' health to merit some form of investment in malaria control [@Pluess2009]. Insecticide fumigation, bednet distribution, test-and-treat clinics, and occasional educational campaigns are common forms of corporate social responsibility to combat malaria. A World Bank study on Nigerian sugarcane cutters found that the simple availability of testing and treatment increased productivity by 10% in the weeks following the provision of services, the conclusion being that both the treatment and the test result were effective in increasing productivity, the latter simply increasing the information available which could influence personal labor allocation decisions [@Dillon2014].

Malaria may disincentivize investment in business in malaria zones. In a vicious cycle, businesses already in malaria zones may not be incentivized to invest in malaria control, perhaps because this is seen as the domain of the public sector [@Curtis2003], or perhaps due to not knowing the extent to which the benefits in health of that control will accrue to the business itself.

The evidence of malaria's negative effects on both health and wealth, at both the individual [@Cole2006] and collective [@McCarthy2000] levels, as well as in both the short [@AsensoOkyere1997] [@Ajani2010] and long [@Hong2011] [@Sachs2002] terms, are amply described in the public health and economics literature [@Phillips98]. That said, very little exists in the literature examining the costs and benefits of malaria control from a private ledger perspective (ie, the point of view a business investor). Unlike a government or individual, a private firm investing in malaria control may be most interested not in its long-term macroeconomic effects, nor its short-term personal health effects, but rather on the impact on productivity (and the extent to which that producitivty's benefits accrue to the firm), absenteeism, and the opportunity costs of expenditures in malaria control. In other words, though the large magnitude of malaria control's benefits are well known, the portion of those benefits accrued by a private firm investing in malaria control is unknown.

This study does not endeavor to expand the current body of knowledge regarding malaria's ill effects on individuals and societies; rather, it aims to provide empirical evidence pertaining to a facet of malaria economics with very little in the literature: malaria control from a private-sector investment perspective. Our study adds to the existing literature by showing the effect of specific malaria control interventions on worker absenteeism, and translating that effect into a return on investment. Unlike previous studies on the effectiveness of malaria control interventions, our's focuses solely on one firm carrying out one intervention, takes advantage of individual-level data, and analyzes results from a ledger perspective.

## Study area

Sugar has been systematically cultivated in Mozambique since the late 1800s. The Incomati Estates company, a small sugarcane processing facility started by a Scotsman on the banks of the Incomati River in 1913, was the first firm to export sugar from Mozambique. Following its purchase by international investors in the 1950s, it (along with the rest of the industry) expanded significantly, exporting to both Europe and the United States. In the late 1960s, a Portuguese family opened the Maragra Açúcar company, while a group of foreign investors started the nearby Marracuene Agrícola Açucareira mill. By the early 1970s, sugar grew to account for greater than 10% of Mozambique's national exports. Nationalized following independence in 1977, the industry's production levels fell from 320,000 annual tons to fewer than 15,000 by 1992. After the end of the civil war, foreign investment revived the sugar industry, and by 2011 production had surpassed its 1972 peak.

@@

The mill of the Maragra Açucar SA (a subsidary of the Illovo sugar company, henceforth referred to as "Maragra") was nationalized in the 1970s (like all other Mozambican mills), went through a period of low production, and then fell completely out of use by 1984. It re-opened in private hands in 1992, and was renovated by a group of international investors in 1998. Today, Maragra accounts for roughly one quarter of Mozambique's overall sugar production (second only to the nearby Xinavane mill run by the Hulett Sugar Tongaat company) [@sutton]. With a favorably close location to the port of Maputo, ample land, approximately 5,000 employees (of which three quarters are seasonal), and a mill with the capacity to process not only all the sugarcane grown on Maragra's land, but also the cane of the many nearby smallholders [@OLaughlin2016], Maragra has so far been able to weather the 2016 Mozambican crisis and concurrent collapse in global sugar prices.

Maragra (figure 1, panel iii) is located in the district of Manhiça (figure 1, panel ii), a semi-rural area in the south of Mozambique (figure 1, panel i). 80 kilometers north of the Mozambican capital of Maputo, the district is low-lying, consisting largely of savannah and wetlands along the Incomati River. Most of the areas 160,000 residents [@Sacoor2013] work as subsistence farmers. Migration from the area to South Africa for the purpose of employment in the profitable construction industry is common, especially among men [@Nhacolo_2006], as is migration to the area (from other parts of Mozambique) for seasonal work on the sugarcane plantations at Maragra and the slightly larger facility in Xinavane (at the district's border with Magude) [@OLaughlin2016].

Poverty is rife in southern Mozambique, and its associated illnesses take their toll on the population. The community prevalence of HIV/AIDS is as high as 40% [@Gonzlez2012]; even in a more recent study suggesting a much lower prevalence of 22%, the risk of infection is still twice that of nearby areas [@Mocumbi2017]. Recent years have seen a three-fold increase in tuberculosis [@GarcaBasteiro2017]. Malaria, which has the greatest mortality burden due to the fact that the young are especially vulnerable to its effects, is perennial, though worse during the rainy season (December - March) [@Sate2003]. Adult malaria is essentially universal, albeit much of it asymptomatic [@Mayor2007]. In regards to worker health, malaria is Maragra's primary concern, being so important as to justify the existence of both an on-site testing laboratory and clinic, as well as a malaria control department.

Maragra workers are mostly seasonal, working for the firm approximately half of the year during harvest time, and cultivating crops, working in construction, or going unemployed (or working elsewhere) during the other half. Though many workers live "on-site" (ie, within the delineated property of the firm itself), a sizeable minority resides in the environs (figure 1, panel iv). Maragra provides indoor residual spraying (IRS) using ACT (alpha-cypermethrin) and DDT (Dichlorodiphenyltrichloroethane) depending on stocks (a preference apparently exists for the former, but it is not always available). IRS activities, managed by Maragra's Malaria Control program, are ongoing throughout the year (Figure 1, panel i). Nearly all on-site houses are sprayed, though the time between fumigations is irregular (Figure 2, panel ii). Off-site houses may also receive IRS (managed and carried out by the National Malaria Control Program), but the status and timing of these fumigations is not known at the individual level.

@@

>>>>

@@

```{r}
# Calculate area of maragra
# x <- bairros
# x$NeighCode <- as.numeric(as.character(x$NeighCode))
# x@data <- x@data %>%
#   mutate(maragra_bairro = NeighCode %in%
#            c(601:609,701:709, 1001:1003)) %>%
#   mutate(maragra_fabrica = NeighCode %in%
#            1001:1003)
# x <- spTransform(x,CRS("+proj=utm +zone=36 +datum=WGS84 +units=m") )
# 
# the_bairro <- x[x$maragra_bairro,]
# the_fabrica <- x[x$maragra_fabrica,]
# # the_bairro <- spTransform(the_bairro, CRS("+proj=utm +zone=36 +datum=WGS84 +units=m") )
# # the_fabrica <- spTransform(the_fabrica, CRS("+proj=utm +zone=36 +datum=WGS84 +units=m") )
# 
# rgeos::gArea(the_bairro) / 1000000
# rgeos::gArea(the_fabrica) / 1000000

# Used this for plantation estimation: https://www.daftlogic.com/projects-google-maps-area-calculator-tool.htm
```

@@

\newpage

### 2018-05-11 inst/rmd/03_results.Rmd
removed text

We collected absenteeism and demographic data data on `r n` workers from 2012 through 2016. Workers were approximately 60% male, and more than 80% fieldworkers (predominantly cane-cutters). Most were in their 20s and 30s and employed on temporary contracts. Table 1 shows overall worker details divided by "Treatment" versus "Control" status. "Treatment" is considered the time beginning at IRS administration and ending six months later; "Control" is considered time observed either prior to IRS administration (and greater than 6 months after previous IRS administration) or among workers who never received IRS. Because of the temporal nature of these categories, many workers belong to both treatment and control groups, albeit at different times.

@@

    out <- out %>% dplyr::select(variable, ` `, `Treatment workers`, `Treatment days`, `Control workers`, `Control days`)
    out$` ` <- as.character(out$` `)
  return(out)
}
out_list <- list()
vars <- c('department', 
           'sex',
           'status',
           'age')
for(i in 1:length(vars)){
  message(i)
  y <- calculate(var = vars[i], data = x)
  out_list[[i]] <- y
}
demo_table <- bind_rows(out_list)
names(demo_table)[1:2] <- c('Variable', ' ')
print(xtable::xtable(demo_table,
                     caption = 'Overall worker characteristics',
                     align = c('r', 'r', rep('l', 5))),
      include.rownames = FALSE)

@@

Temporary workers tended to be younger and male; female temporary workers being on average 5-10 years older than their male counterparts. Permanent workers had a bi-modal age distribution, the older peak explained by the greater density of males in administrative roles. Females accounted for 43% of temporary workers but only 16% of permanent workers (figure x).

@@

During the study period, weather followed typical seasonal trends for the area, albeit slightly drier than previous periods.

@@

In Southern Mozambique, malaria peaks during the summer months (December through March) most years (Figure 5, panel A), and worker absenteeism at Maragra rates track malaria incidence closely, following the same seasonal patterns (figure 4, panel B). Both all-cause absenteeism and sick absenteeism have declined in recent years at Maragra (figure 5, panel C), with the latter declining at a faster rate than the former. The fact that the rate of confirmed cases at the company clinic is largely non-seasonal (figure 5, panels E and F) suggests that a significant portion of workers either seek care for malaria elsewhere (for example, government health posts, of which several are nearby and in some many cases closer to workers' residence than the company clinic) or do not seek care during malaria infection. Accordingly, we focus our analysis on all-cause absenteeism rather than sickness absenteeism or malaria diagnostics, with the assumption that much of illness is captured by absenteeism but not by the clinical data.

@@

  out <- out %>%
    spread(key = year,
           value = val)
  out$variable <- as.character(out$variable)
  new_var <- ifelse(var == 'on_site',
                    'Residence',
                    ifelse(var == 'rainy',
                           'Precipitation',
                           var))
  left <- data.frame(x = c(new_var, rep('', (nrow(out) - 1))))
  out <- cbind(left, out)
  names(out)[1:2] <- c('Variable', 'x')
  # out$variable[1] <- paste0(Hmisc::capitalize(new_var), ': ', out$variable[1])
  return(out)
}
out_list <- list()
vars <- c('season', 
           'field',
          'permanent_or_temporary',
           'sex',
           'on_site',
           'rainy')
for(i in 1:length(vars)){
  x <- calculate(var = vars[i])
  out_list[[i]] <- x
}
ab_table <- bind_rows(out_list)
ab_table$Variable <- Hmisc::capitalize(ab_table$Variable)
ab_table$Variable[ab_table$Variable == 'Season'] <- 'Malaria season'
ab_table$Variable[ab_table$Variable == 'Field'] <- 'Worker type'
ab_table$Variable[ab_table$Variable == 'Permanent_or_temporary'] <- 'Contract'
ab_table$x <- Hmisc::capitalize(ab_table$x)
names(ab_table)[2] <- ''
print(xtable::xtable(ab_table,
                     caption = 'Absenteeism rate by year and worker characteristics',
                     align = c('r', 'r', 'l', rep('l', 4))),
      include.rownames = FALSE)
```

@@

The extent to which rainfall's effect on absenteeism is confounded by its effect on malaria is discussed in the model results section.

**Costs of program**: The malaria control program at Maragra has an annual operating budget of approximately $112,000, which includes the purchase of insecticide, the wages of IRS sprayers and drivers, transportation, record-keeping, and general administrative costs. Assuming linearity in costs, the program spends approximately $19 per building sprayed. With each agregado containing an average of 2.2 workers. Much of the benefit of IRS goes to non-worker residents of sprayed agregados (who constitute a majority), but this benefit is purposefully ignored for this analysis.

**Cost of malaria**: Given the likelihood that clinical data does not fully capture all malaria cases (and most likely captures only a small fraction of actual infections, given the high rates of acumulated adult immunity [Mayor2007]), we do not quantify the costs of malaria infection to the company. Rather, we first estimate the reduction in absenteeism attributible to IRS, and then quantify the savings associated with prevented absences. Additionally, we calculate the clinical savings of IRS by first estimating the share of absences which are associated with an episode of clinical malaria, and then applying the clinical cost per case to the equivalent share of prevented absences. We intentionally ignore the savings accrued by the public health system, as well as the likely utility gains in secondary realms such as school absenteeism, producitivity, etc.

## Analysis

### Effect of IRS on absenteeism

Immediately following IRS, a worker's likelihood of absence drops significantly (appendix figure, panel A). As one would expect if the mechanism by which IRS reduces absence is through reduced malaria infection, the effect of IRS during the low transmission season is significant, but far less substantial in effect size (appendix figure, panel B).

In order to assess the effect of IRS on absenteeism, we created 4 worker fixed effects models for the 4 principal worker types (permanent field worker, temporary field worker, permanent non-field worker, temporary non-field worker) for both outcoms (all cause absenteeism vs. sickness absenteeism). The reason for segregating models rather than incorporating worker type as a fixed effect was that it seemed plausible that the effect of the treatment (IRS) on the outcome would be different (and vary differently over time) depending on these worker types. For instance, it is reasonable to assume that IRS' effect would be differential for a temporary worker (who likely lives in the fumigated house less than 100% of the year) than for a permanent worker.

Given that IRS is carried out on a continuous rolling basis, we used relative worker time (as opposed to calendar) as the time component of our model - that is, each workers' days until and since IRS were standardized so that day 0 (date of fumigation) was the interruption point. This has the advantage of incorporating into the model a variety of potential time-varying confounders in a quasi-randomized fashion, thereby making it unecessary to adjust for them specifically. Our model took into account workers' residential location (on or off site) - even though off site workers by definition could not be administered IRS by Maragra - so as to capture other seasonal effects from those workers.

The herd protection score, unit-less by nature given that it's a summation of binary protection statuses weighted by inverse-transformed distances, took the following distribution in our data:

@@

The model results for all cause absenteeism (table 4) suggest a significant decrease in absenteeism following the administration of IRS. The effect is ambivalent among temporary workers, whereas among permanent workers the administration of IRS decreases absenteeism by xxx. These effects are consistent with IRS decreasing absence through the medium of malaria, since (a) the effect is not as strong in the first month (potentially during the parasite incubation phase), the effect is stronger on field workers (who are more socioeconomically the demographic at greater risk of malaria than their administrative counterparts), and the effect is ambivalent among temporary workers (since the protective effect of IRS is only a function of the amount of time that the worker spends in the fumigated house).

@@

```{r}
# Make dummy data for seeing effect of herd on irs
```

@@

## Translating absences to costs

xxx\todo{Not done yet. This will be relatively straightforward arithmetic, but I want to first make sure that the models are rock solid (since all of the "translation" hinges on them).}

## Return on investment

\todo{Not done yet. In conversation with Eduardo from MC about getting more accurate costs}

**Savings**

- In percentage point terms, reduction from 13% to 8%.  
- 5 annually prevented absences per worker.   
- 8,000 workers: 40,000 prevented absences, wage of 3 USD
- TOTAL: 120,000 USD in productivity-only savings.

**Costs**

- 8 IRS workers, 1500 USD yearly = 12,000 USD  
- ACT + DDT: 50,000 USD 
- Facilities, vehicles, gas: 50,000 USD  
- TOTAL: 112,000 USD in IRS-only costs

**7% ROI** (ignoring clinical costs)

@@

Two principal concerns call into question the results of our analysis. First, the application of IRS to a workers house may be endogenous.\todo{Is this the right term here? We're not dealing with an "unobserved" factor per se, rather we're dealing with a situation in which the cause might influence the effect, and the effect might influence the cause, ie a feedback loop.} It is reasonable to suspect that the application of IRS to households is not random, but rather that IRS was applied more frequently to houses which had already seen a malaria case. In this case, our estimated effect of IRS on absenteeism would likely be underestimated, with the post-IRS absenteeism rates actually having declined from a greater pre-IRS absenteeism rate than otherwise suggested.

To check for this, we estimate the odds of absenteeism as a function of receiving IRS _during the 10 day period prior to receiving IRS_. If IRS application is indeed endogenous, we would expect absenteeism to be elevated during this period (since the increase in absenteeism would be theoretically responsible for the application of IRS), a situation which would require further statistical adjustment. If, on the other hand, there is no endogeneity, we would expect absenteeism in the 10 day period prior to IRS administration to be similar to other pre-IRS absenteeism (adjusting for seasonality).\todo{Should add model specification formula here.}

The below shows our robustness check for all cause absenteeism.

@@

make_model <- function(formula = 'absent ~ pre_irs + time_period'){
  ro_model <- glm(as.formula(formula), 
                data = ro_data,
                family = binomial('logit'))
  coefs <- exp(coef(ro_model))
  ci <- confint(ro_model)
  ci <- exp(ci)
  out <- data.frame(Variable = names(coefs),
                    Estimate = coefs,
                    Lower = ci[,1],
                    Upper = ci[,2])
  row.names(out) <- NULL
  out$`P value` <- coef(summary(ro_model))[,4]
  out$Significant <- ifelse(out$`P value` <= 0.001, 'xxx',
                            ifelse(out$`P value` <= 0.01, 'xx',
                                   ifelse(out$`P value` <= 0.05, 'x',
                                          '')))
  out$Variable <- as.character(out$Variable)
  out$Variable <- gsub('pre_irsTRUE', '10 days prior to IRS', out$Variable)
  out$Variable <- gsub('time_period', '', out$Variable)
  return(kable(out, format = "latex") %>%
  kable_styling() %>%
    row_spec(row = which(grepl('prior', out$Variable)),background = "yellow"))
}
```

>>>>

We collected absenteeism and demographic data data on `r n` workers from 2012 through 2016.

@@

    out <- out %>% dplyr::select(variable, ` `, `Treatment workers`, `Treatment days`, `Control workers`, `Control days`)
    out$` ` <- as.character(out$` `)
  return(out)
}
out_list <- list()
vars <- c('department', 
           'sex',
           'status',
           'age')
for(i in 1:length(vars)){
  message(i)
  y <- calculate(var = vars[i], data = x)
  out_list[[i]] <- y
}
demo_table <- bind_rows(out_list)
names(demo_table)[1:2] <- c('Variable', ' ')
kable(demo_table, format = 'html') %>% kable_styling()
# print(xtable::xtable(demo_table,
#                      caption = 'Overall worker characteristics',
#                      align = c('r', 'r', rep('l', 5))),
#       include.rownames = FALSE)

@@

@@

@@

@@

  out <- out %>%
    spread(key = year,
           value = val)
  out$variable <- as.character(out$variable)
  new_var <- ifelse(var == 'on_site',
                    'Residence',
                    ifelse(var == 'rainy',
                           'Precipitation',
                           var))
  left <- data.frame(x = c(new_var, rep('', (nrow(out) - 1))))
  out <- cbind(left, out)
  names(out)[1:2] <- c('Variable', 'x')
  # out$variable[1] <- paste0(Hmisc::capitalize(new_var), ': ', out$variable[1])
  return(out)
}
out_list <- list()
vars <- c('season', 
           'field',
          'permanent_or_temporary',
           'sex',
           'on_site',
           'rainy')
for(i in 1:length(vars)){
  x <- calculate(var = vars[i])
  out_list[[i]] <- x
}
ab_table <- bind_rows(out_list)
ab_table$Variable <- Hmisc::capitalize(ab_table$Variable)
ab_table$Variable[ab_table$Variable == 'Season'] <- 'Malaria season'
ab_table$Variable[ab_table$Variable == 'Field'] <- 'Worker type'
ab_table$Variable[ab_table$Variable == 'Permanent_or_temporary'] <- 'Contract'
ab_table$x <- Hmisc::capitalize(ab_table$x)
names(ab_table)[2] <- ''
kable(ab_table, format = 'html') %>% kable_styling()
# print(xtable::xtable(ab_table,
#                      caption = 'Absenteeism rate by year and worker characteristics',
#                      align = c('r', 'r', 'l', rep('l', 4))),
#       include.rownames = FALSE)
```

@@

@@

@@

@@

@@

@@

make_model <- function(formula = 'absent ~ pre_irs + time_period'){
  ro_model <- glm(as.formula(formula), 
                data = ro_data,
                family = binomial('logit'))
  coefs <- exp(coef(ro_model))
  ci <- confint(ro_model)
  ci <- exp(ci)
  out <- data.frame(Variable = names(coefs),
                    Estimate = coefs,
                    Lower = ci[,1],
                    Upper = ci[,2])
  row.names(out) <- NULL
  out$`P value` <- coef(summary(ro_model))[,4]
  out$Significant <- ifelse(out$`P value` <= 0.001, 'xxx',
                            ifelse(out$`P value` <= 0.01, 'xx',
                                   ifelse(out$`P value` <= 0.05, 'x',
                                          '')))
  out$Variable <- as.character(out$Variable)
  out$Variable <- gsub('pre_irsTRUE', '10 days prior to IRS', out$Variable)
  out$Variable <- gsub('time_period', '', out$Variable)
  return(kable(out, format = "html") %>%
  kable_styling() %>%
    row_spec(row = which(grepl('prior', out$Variable)),background = "yellow"))
}
```

### 2018-05-11 inst/rmd/paper.Rmd
removed text

---
title: Is malaria control profitable? Return on investment of privately-managed residential fumigations at a large sugarcane processing facility in Southern Mozambique
date: ''
output: 
  pdf_document:
    keep_tex: true
    latex_engine: xelatex
    includes:
      in_header: preamble.sty
fig_width: 4
fig_height: 2.6
bibliography: bibliography.bib
csl: journal-of-health-economics.csl
---

@@

# Packages
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(maragra)
library(knitr)
library(RColorBrewer)
library(extrafont)
library(kableExtra)
loadfonts()
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               dev = "cairo_pdf",
               fig.pos="!h",
               fig.path = 'figures/')
opts_knit$set(width=75)
options(xtable.comment = FALSE)

>>>>

---
title: Is malaria control profitable? Return on investment of privately-managed residential
  fumigations at a large sugarcane processing facility in Southern Mozambique
csl: apa_modified.csl
date: ''
output:
  html_document:
    df_print: paged
  pdf_document:
    includes:
      in_header: preamble.sty
    keep_tex: yes
    latex_engine: xelatex
  word_document: default
fig_width: 4
fig_height: 2.6
bibliography: bibliography.bib
---

@@

# Packages
library(tidyverse)
library(knitr)
library(Hmisc)
library(brew)
library(maragra)
library(knitr)
library(RColorBrewer)
library(extrafont)
library(kableExtra)
loadfonts()
## Global options
options(max.print="75")
opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               # dev = "cairo_pdf",
               fig.pos="!h",
               fig.path = 'figures/')
opts_knit$set(width=75)
options(xtable.comment = FALSE)
